---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# sassr

Sassr is a CSS preprocessor, letting you use variables, nesting, and inheritance in CSS.

Sassr uses the [Sass](https://sass-lang.com/) CSS extension language. Sass is stable, powerful, and CSS compatiable. Sassr is an R wrapper for [LibSass](https://github.com/sass/libsass), a fast Sass compiler written in C++.

## Installation

You can install the released version of sass from [CRAN](https://CRAN.R-project.org) with:

```r
install.packages("sass")
```

You can install the latest development build from Github with:

```r
# install.packages("devtools")
devtools::install_github("tmastny/sass")
```

## Example

At it's most basic, Sass can do computations with standard CSS code:

```{r}
library(sassr)
compile_sass(text = "foo { margin: 122px * .3; }")
```

For an overview of the major features such as variables, nesting, and imports check out the [Sass Basics](https://sass-lang.com/guide).

