---
title: "Introduction to Sass"
author: "Timothy Mastny"
date: "2018-08-14"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

For single [Rmarkdown](https://rmarkdown.rstudio.com/) documents or [Shiny](http://shiny.rstudio.com/) apps, styling with CSS or using custom themes such as [shinythemes](https://github.com/rstudio/shinythemes) works well.

As an R developer, designing a default style sheet is difficult. The developer wants to provide a theme that works and looks good, while providing customization options for users.

This article will briefly show how to use `sass` to generate style sheets based on user input.

## Variables

Sass variables are key to generating dynamic CSS. Because `sass` compiles fast, you can define components of your styling at runtime based on user input.

It is best practice to define your [variables](http://sass-lang.com/guide#topic-2) separately, and then [import](http://sass-lang.com/guide#topic-5) them into the main Sass file. For example, we'll define a color variable and use it to set the body's background color.

```css
/* _variables.scss */
$color: #FFFFFF;
```
```css
/* new-style.scss */
@import 'variables';
body {
  background-color: $color;
}
```

The following [shiny app]("https://gallery.shinyapps.io/138-sass-color") will demonstrate this example. We'll use Dean Attali's [colourpicker](https://github.com/daattali/colourpicker) to choose a color and write it to `_variables.scss`.

Then the app compiles `new-style.scss` and renders the new style.

```{r echo = FALSE}
knitr::include_app("https://gallery.shinyapps.io/138-sass-color", height = "600px")
```

## Functions

If we choose a background color that is too dark, we won't be able to read the title **Sass Example**. Therefore, we need to make the font color depend on the background color.

Sass offers many [built-in functions](http://sass-lang.com/documentation/file.SASS_REFERENCE.html#functions) that help with dynamic styling, and you can always write [your own](http://sass-lang.com/documentation/file.SASS_REFERENCE.html#function_directives).

```css
/* _variables.scss */
$color: #FFFFFF;
```
```css
/* new-style.scss */
@import 'variables';
body {
  background-color: $color;
}

/* https://stackoverflow.com/a/3943023/6637133 */
@function font-color($color) {
  @return if(
    red($color) * 0.299 + green($color) * 0.587 + blue($color) * 0.114 > 186,
    #000000, #ffffff
  );
}

h1 {
  color: font-color($color);
}
```

In our next [shiny app](https://gallery.shinyapps.io/139-sass-font/), the font color now [depends](https://stackoverflow.com/a/3943023/6637133) on the background color, switching between black and white as necessary.

```{r echo = FALSE}
knitr::include_app("https://gallery.shinyapps.io/139-sass-font/", height = "400px")
```

## Resizing

Dynamic resizing based on user-input is also useful. Sass is fully CSS compatible, so any styling option is available.

In our last [shiny app](https://gallery.shinyapps.io/140-sass-size/) we'll change the width of the image based on a slider:

```css
/* _variables.scss */
$color: #FFFFFF;
$width: 100;
```
```css
/* new-style.scss */
@import 'variables';
body {
  background-color: $color;
}

/* https://stackoverflow.com/a/3943023/6637133 */
@function font-color($color) {
  @return if(
    red($color) * 0.299 + green($color) * 0.587 + blue($color) * 0.114 > 186,
    #000000, #ffffff
  );
}

h1 {
  color: font-color($color);
}

.shiny-plot-output {
  max-width: percentage($width / 100);
}
```

```{r echo = FALSE}
knitr::include_app("https://gallery.shinyapps.io/140-sass-size/", height = "600px")
```

## CSS Formatting

Sass can also produce human readable or production ready CSS, depending on the need.

Here we'll convert the human readable Sass to minified CSS.

```{r}
new_style <- "
  $color: #FFFFFF;
  $width: 100;

  body {
    background-color: $color;
  }

  /* https://stackoverflow.com/a/3943023/6637133 */
  @function font-color($color) {
    @return if(
      red($color) * 0.299 + green($color) * 0.587 + blue($color) * 0.114 > 186,
      #000000, #ffffff
    );
  }

  h1 {
    color: font-color($color);
  }

  .shiny-plot-output {
    max-width: percentage($width / 100);
  }
"
```

```{r}
library(sass)
opts <- sass_options(output_style = "compressed")
sass::sass(new_style, options = opts)
```
